


# Basic analysis of UW Archaeology PhD theses

There are no clear guidelines about the length or structure of a PhD thesis in 
archaeology at UW. To answer these questions, we decided to make a quick study
of the norms evident in PhD theses produced in the last ten years. 

## Methods

We made counts of the total number of pages for each theses and the number of pages per chapter. We also made a note of the year the thesis was passed to examine trends over time. We entered the data into a google sheet.

```{r echo = T, message = F, cache = F}
# setup for interactive plots
require(rCharts)
options(RCHART_WIDTH = 600, RCHART_HEIGHT = 400)
knitr::opts_chunk$set(comment = NA, results = 'asis', 
                      tidy = F, message = F)
```


```{r message=FALSE, warning=FALSE, cache=TRUE, tidy=FALSE}
# get data from google sheet
# connect to google sheet
require(RCurl) # if you get an error here, see these
# http://www.dummies.com/how-to/content/
# how-to-install-load-and-unload-packages-in-r.html
# http://www.statmethods.net/interface/packages.html
# http://www.youtube.com/watch?v=u1r5XTqrCTQ
options(RCurlOptions = list(capath = system.file("CurlSSL", 
                                                 "cacert.pem", 
                                                 package = "RCurl"), 
                            ssl.verifypeer = FALSE))
#in google spreadsheet, go to file-> publish to web -> 
# get link to publish to web -> get csv file
goog <- "https://docs.google.com/spreadsheet/pub?key=0As7CmPqGXTzldC1yT2ZHUURDN1o1a0pBdndxVVFyNHc&output=csv"
data <- read.csv(textConnection(getURL(goog)), 
                 stringsAsFactors = FALSE)
```


Has the length of a PhD thesis changed over time? Look at total pages...
```{r message=FALSE, warning=FALSE}
# plot year and total pages
require(ggplot2)
ggplot(data, aes(year, total.pages)) +
  # geom_point() +
  geom_text(aes(label = author.surname)) +
  geom_rug(position = position_jitter(w = 0.1)) + 
 # geom_smooth() +
  theme_minimal() +
  xlab("Year of thesis submission") +
  ylab("Total number of pages")
```

Has the length of a PhD thesis changed over time? Look at body pages...
```{r message=FALSE, warning=FALSE}
# plot year and body pages
require(ggplot2)
ggplot(data, aes(year, pages.of.body)) +
  # geom_smooth() +
  # geom_point() +
  geom_text(aes(label = author.surname)) +
  geom_rug(position = position_jitter(w = 0.1)) + 
  theme_minimal()
```


How many chapters does a PhD thesis have?
```{r message=FALSE, warning=FALSE, tidy=FALSE}
require(reshape2)
data_m <- melt(data, id.var = c('year',  'author.surname' ))
data_ch <- data_m[data_m$variable %in% 
                    sapply(1:10, function(i) paste0('pages.ch',i)), ]
# get table of count of thesis having each chapter
tab_ch <- aggregate(value ~ variable, data_ch, length) 
# make new numeric variable for plotting - extract numbers only
tab_ch$chapter <- as.numeric(gsub("\\D", "", tab_ch$variable))

# plot histogram of chapter distribution
ggplot(tab_ch, aes(chapter, value)) +
  geom_bar(stat="identity") +
  theme_minimal() +
  scale_x_continuous(breaks=1:10)
```


What is the relationship between the number of chapters and the number of pages?
```{r message=FALSE, warning=FALSE, results = 'asis', comment = NA, message = F}
# plot number of chapters and total pages
require(ggplot2)
ggplot(data, aes(number.of.chapters, total.pages)) +
  geom_smooth(se = FALSE) +
  # geom_point() +
  geom_text(aes(label = author.surname)) +
  geom_rug(position = position_jitter(w = 0.1)) + 
  theme_minimal()
```

```{r chart1}
# interactive
require(rCharts) 
# javascript uses periods for indexing...
data$authorsurname    <- data$author.surname 
data$totalpages       <- data$total.pages
data$numberofchapters <- data$number.of.chapters
n1 <- nPlot(totalpages ~ numberofchapters , data = data, type = "scatterChart")
# n1$addParams(width = 600, height = 300)
n1$chart(
  showDistX = TRUE,
  showDistY = TRUE,
  tooltipContent = "#! function(key, x, y, e){ 
  return '<b>Author</b>: ' + e.point.authorsurname
  } !#"
)
n1$print('chart1')
```


How many pages in each chapter?
```{r message=FALSE, warning=FALSE}
# boxplot
require(scales)
ggplot(data_ch, aes(variable, as.numeric(value))) +
  geom_boxplot(notch = FALSE) + # if notches overlap = no sig diff
  # geom_violin() +
  theme_minimal() +
  # scale_y_continuous(trans = log2_trans()) +
  # scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") 

# plot with authors as data labels and smoother
ggplot(data_ch, aes(variable, as.numeric(value))) +
  geom_text(aes(label = author.surname, size = 0.1)) +
  # geom_smooth(aes(group = 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") 
```

```{r message=FALSE, warning=FALSE, echo=FALSE, eval = FALSE}
knitr::knit2html("uw_archy_phd_theses.rmd")
system(paste0("pandoc -o ", 'uw_archy_phd_theses', ".pdf ", 'uw_archy_phd_theses', ".md"))

filen <- 'uw_archy_phd_theses' # name of the markdown file without suffix
  knitr::knit(paste0(filen,".Rmd"))
# make PDF
  system(paste0("pandoc -s ", paste0(filen,".md"), " -t latex -o ", paste0(filen,".pdf"), " --highlight-style=tango  -S"))
# make DOCX
  system(paste0("pandoc -s ", paste0(filen,".md"), "  -o ", paste0(filen,".docx"), " --highlight-style=tango  -S"))


# summary stats for chapters
# aggregate(as.numeric(data_ch$value) ~ data_ch$variable, FUN = "summary")
```
  



